---
title: "Journal"
author: "Mark Moawad"
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Welcome and GROOTings

This is my Journal where I will be publishing my progress in Data Science Module courses at TUHH!

## Welcome

Myself in brief, I am Mark Moawad; a car guy, roboticist and hobbyist!
I come from the great country of Pharaohs; Egypt!
I love dogs, LEGO, motorcycles and music. If you share some or many interests with me don't hesitate to reach out, we might be good friends at the end of the day!

## GROOTings

I meant to say "GROOTings" instead of Greetings, because I am a huge fan of GROOT, actually I am GROOT, you are GROOT too. We are all GROOT!
Have a GROOT day!


# Introduction (Chapter 2)

## Goal of Data Science

A Data Scientist's main job is to study the products, look for opportunities to sell new products, better serve the customer and better market the products. All of that is supposed to be justified by data.

## Goal of this Analysis

The goal of this case study is to analyze the sales of bikes sold through bike stores in Germany. The study is conducted many times, each according to different parameters to get meaningful results. That will be done by importing, wrangling and visualizing of the provided data.

### Studies

#### Sales by year
#### Sales by year and Main bike Category
#### Sales by location
#### Sales by location and year


#### Sales by year results

```{r code, results='hide'}
# 1.0 Load libraries ----
library(tidyverse)

library(readxl)

# 2.0 Importing Files ----
bikes_tbl      <- read_excel(path = "DS_101/00_data/01_bike_sales/01_raw_data/bikes.xlsx")

orderlines_tbl <- read_excel("DS_101/00_data/01_bike_sales/01_raw_data/orderlines.xlsx")

bikeshops_tbl  <- read_excel("DS_101/00_data/01_bike_sales/01_raw_data/bikeshops.xlsx")

# 3.0 Joining Data ----

bike_orderlines_joined_tbl <- orderlines_tbl %>%
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))


# 4.0 Wrangling Data ----
bike_orderlines_joined_tbl %>% 
  select(category) %>%
  filter(str_detect(category, "^Mountain")) %>% 
  unique()

bike_orderlines_wrangled_tbl <- bike_orderlines_joined_tbl %>%
# 4.1 Separate category name
separate(col    = category,
         into   = c("category.1", "category.2", "category.3"),
         sep    = " - ") %>%
  
# 4.2 Add the total price (price * quantity) 
# Add a column to a tibble that uses a formula-style calculation of other columns
mutate(total.price = price * quantity) %>%

# 4.3 Optional: Reorganize. Using select to grab or remove unnecessary columns
# 4.3.1 by exact column name
select(-...1, -gender) %>%

# 4.3.2 by a pattern
# You can use the select_helpers to define patterns. 
# Type ?ends_with and click on Select helpers in the documentation
select(-ends_with(".id")) %>%

# 4.3.3 Actually we need the column "order.id". Let's bind it back to the data
bind_cols(bike_orderlines_joined_tbl %>% select(order.id)) %>% 

# 4.3.4 You can reorder the data by selecting the columns in your desired order.
# You can use select_helpers like contains() or everything()
select(order.id, contains("order"), contains("model"), contains("category"),
       price, quantity, total.price,
       everything()) %>%

# 4.4 Rename columns because we actually wanted underscores instead of the dots
# (one at the time vs. multiple at once)
rename(bikeshop = name) %>%
set_names(names(.) %>% str_replace_all("\\.", "_"))

# 5.0 Business Insights ----
# 5.1 Sales by Year ----

library(lubridate)
# Step 1 - Manipulate
sales_by_year_tbl <- bike_orderlines_wrangled_tbl %>%
  
# Select columns
select(order_date, total_price) %>%
  
# Add year column
mutate(year = year(order_date)) %>%
 
# Grouping by year and summarizing sales
group_by(year) %>% 
summarize(sales = sum(total_price)) %>%

# Optional: Add a column that turns the numbers into a currency format 
# (makes it in the plot optically more appealing)
# mutate(sales_text = scales::dollar(sales)) <- Works for dollar values
mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                   decimal.mark = ",", 
                                   prefix = "", 
                                   suffix = " €"))


# Step 2 - Visualize
sales_by_year_tbl %>%
  
# Setup canvas with the columns year (x-axis) and sales (y-axis)
ggplot(aes(x = year, y = sales)) +

# Geometries
geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
geom_label(aes(label = sales_text)) + # Adding labels to the bars
geom_smooth(method = "lm", se = FALSE) + # Adding a trendline

# Formatting
# scale_y_continuous(labels = scales::dollar) + # Change the y-axis. 
# Again, we have to adjust it for euro values
scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                  decimal.mark = ",", 
                                                  prefix = "", 
                                                  suffix = " €")) +
labs(
  title    = "Revenue by year",
  subtitle = "Upward Trend",
  x = "", # Override defaults for x and y
  y = "Revenue"
)
```

#### Sales by year and Main bike Category results

```{r}
# 5.2 Sales by Year and Category ----

# Step 1 - Manipulate
sales_by_year_cat_1_tbl <- bike_orderlines_wrangled_tbl %>%
  
# Select columns and add a year
select(order_date, total_price, category_1) %>%
mutate(year = year(order_date)) %>%

# Group by and summarize year and main category
group_by(year, category_1) %>%
summarise(sales = sum(total_price)) %>%
ungroup() %>%

# Format $ Text
mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                   decimal.mark = ",", 
                                   prefix = "", 
                                   suffix = " €"))

# Step 2 - Visualize
sales_by_year_cat_1_tbl %>%
  
# Set up x, y, fill
ggplot(aes(x = year, y = sales, fill = category_1)) +

# Geometries
geom_col() + # Run up to here to get a stacked bar plot

# Facet
facet_wrap(~ category_1) +

# Formatting
scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                  decimal.mark = ",", 
                                                  prefix = "", 
                                                  suffix = " €")) +
labs(
  title = "Revenue by year and main category",
  subtitle = "Each product category has an upward trend",
  fill = "Main category" # Changes the legend name
)
```

#### Sales by location results

```{r fig.width=9, fig.height=7}

# 1.0 Importing Files ----
bikes_tbl_challenge <- read_excel(path = "DS_101/00_data/01_bike_sales/01_raw_data/bikes.xlsx")

orderlines_tbl_challenge <- read_excel("DS_101/00_data/01_bike_sales/01_raw_data/orderlines.xlsx")

bikeshops_tbl_challenge <- read_excel("DS_101/00_data/01_bike_sales/01_raw_data/bikeshops.xlsx")


# 2.0 Joining Data ----

bike_orderlines_joined_tbl_challenge <- orderlines_tbl_challenge %>%
  left_join(bikes_tbl_challenge, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl_challenge, by = c("customer.id" = "bikeshop.id"))

# 3.0 Wrangling Data ----

bike_orderlines_wrangled_tbl_challenge <- bike_orderlines_joined_tbl_challenge %>%

# 3.1 Separate category name
separate(col    = location,
         into   = c("city", "state"),
         sep    = ", ") %>%
  
# 3.2 Add the total price (price * quantity) 
mutate(total.price = price * quantity) %>%

select(-...1, -gender) %>%

select(-ends_with(".id")) %>%

bind_cols(bike_orderlines_joined_tbl_challenge %>% select(order.id)) %>% 

select(order.id, contains("order"), contains("model"), contains("category"),
       price, quantity, total.price,
       everything()) %>%

rename(bikeshop = name) %>%
set_names(names(.) %>% str_replace_all("\\.", "_"))

# 4.0 Business Insights ----
# 4.1 Sales by Location ----

library(lubridate)
# Step 1 - Manipulate
sales_by_location_tbl <- bike_orderlines_wrangled_tbl_challenge %>%
  
# Select columns
select(state, total_price) %>%

# Grouping by state and summarizing sales
group_by(state) %>% 
summarize(sales = sum(total_price)) %>%

# Optional: Add a column that turns the numbers into a currency format 
# (makes it in the plot optically more appealing)
# mutate(sales_text = scales::dollar(sales)) <- Works for dollar values
mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                   decimal.mark = ",", 
                                   prefix = "", 
                                   suffix = " €"))


# Step 2 - Visualize
sales_by_location_tbl %>%
  
# Setup canvas with the columns state (x-axis) and sales (y-axis)
ggplot(aes(x = state, y = sales)) +

# Geometries
geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
geom_label(aes(label = sales_text)) + # Adding labels to the bars
geom_smooth(method = "lm", se = FALSE) + # Adding a trendline

# Formatting
# scale_y_continuous(labels = scales::dollar) + # Change the y-axis. 
# Again, we have to adjust it for euro values
scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                  decimal.mark = ",", 
                                                  prefix = "", 
                                                  suffix = " €")) +
labs(
  title    = "Revenue by location",
  subtitle = "",
  x = "state", # Override defaults for x and y
  y = "Revenue"
) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#### Sales by location and year results

```{r plot, fig.width=12, fig.height=9}

# 5.2 Sales by Location and Year ----

# Step 1 - Manipulate
sales_by_location_and_year_tbl <- bike_orderlines_wrangled_tbl_challenge %>%
  
# Select columns and add a year
select(order_date, total_price, state) %>%
mutate(year = year(order_date)) %>%

# Group by and summarize year and state
group_by(year, state) %>%
summarize(sales = sum(total_price)) %>%
ungroup() %>%

# Format $ Text
mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                   decimal.mark = ",", 
                                   prefix = "", 
                                   suffix = " €"))

# Step 2 - Visualize
sales_by_location_and_year_tbl %>%
  
# Set up x, y, fill
ggplot(aes(x = year, y = sales, fill = state)) +

# Geometries
geom_col() + # Run up to here to get a stacked bar plot

# Facet
facet_wrap(~ state) +

# Formatting
scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                  decimal.mark = ",", 
                                                  prefix = "", 
                                                  suffix = " €")) +
labs(
  title = "Revenue by location and year",
  subtitle = "",
  fill = "" # Changes the legend name
) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Web Data Scraping (Chapter 3)

## Challenge 1 results

In this section, the aim was to learn to scrap the web for data using an API. After a lot of research Accuweather API was chosen to acquire the info of 100 Countries.

```{r}
library(httr)
library(jsonlite)
library(dplyr)
library(glue)
my_apikey <- Sys.getenv("pwd")
url <- modify_url(url = "http://dataservice.accuweather.com", path = glue("/locations/v1/topcities/100?apikey={my_apikey}&language=en-us&details=false"))
resp <- GET(url)
country_list <- resp %>% .$content %>% rawToChar() %>% fromJSON()
City <- country_list$EnglishName
Country <- country_list$Country$EnglishName
Region <- country_list$Region$EnglishName
Longitude <- country_list$GeoPosition$Longitude
Latitude <- country_list$GeoPosition$Latitude
Timezone <- country_list$TimeZone$Name
#Optional to add GMT_Offset, it's only removed so that printing fits the screen
#GMT_Offset <- country_list$TimeZone$GmtOffset
country_list_as_df <- data.frame(City, Country, Region,
                                  Longitude, Latitude, Timezone)
head(country_list_as_df, 10)
```

## Challenge 2 results

The goal of this challenge was to scrap the web manually using selector gadget to get the price details of one bike category of ROSE bike manufacturer. A graph was added at the end to give an insight on the price difference between the different MTB bike models of ROSE. The removed elements from the graph are not yet available for sale (coming soon).

```{r}
library(tidyverse) # Main Package - Loads dplyr, purrr, etc.
library(rvest)     # HTML Hacking & Web Scraping
library(xopen)     # Quickly opening URLs
library(jsonlite)  # converts JSON files to R objects
library(glue)      # concatenate strings
library(stringi)   # character string/text processing
library(ggplot2)
url_home <- "https://www.rosebikes.de/fahrr%C3%A4der/mtb"

html_home <- read_html(url_home)

bike_name_tbl_challenge <- html_home %>%
  
  html_nodes('.catalog-category-bikes__title-text') %>%
  
  html_text %>%
  
  str_remove_all("\n") %>%
  
  enframe(name = "position", value = "name")

bike_price_tbl_challenge <- html_home %>%
  
  html_nodes('.catalog-category-bikes__price-title') %>%
  
  html_text %>%
  
  str_remove_all("\n") %>%
  
  enframe(name = "position", value = "price") %>%
  
  na_if("") %>%
  
  mutate(price = price %>% str_remove_all("ab ")) %>%
  
  mutate(price_in_EUR = price %>% str_remove_all("€")) %>%
  
  select(-price) %>% mutate(price_in_EUR = readr::parse_number(.$price_in_EUR))

bike_name_price_tbl <- left_join(bike_name_tbl_challenge, bike_price_tbl_challenge) %>%
  select(-position)
bike_name_price_tbl
ggplot(bike_name_price_tbl, aes(x = name, y = price_in_EUR, color = name)) +
  geom_col() +
  expand_limits(x = 0, y = 0) +
  labs(title = "Bike Model vs Price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Data Wrangling (Chapter 4)
In this chapter, the acquired skills of Data Wrangling were used as much as possible to answer the following questions with Data Analysis:

```{r}
# Importing library: ---- 
library(vroom)
library(tidyverse)
library(data.table)
library(tictoc)

# 2.0 Importing Data ----

# Patents: ----

col_types <- list(
  id = col_character(),
  date = col_date("%Y-%m-%d"),
  num_claims = col_double()
)

patent_tbl <- vroom(
  file       = "DS_101/02_data_wrangling/Patent_data_reduced/patent.tsv",
  delim      = "\t",
  col_types  = col_types,
  na         = c("", "NA", "NULL")
)

#Assignee_id = id,
# Assignee: ----

col_types_assignee <- list(
  id = col_character(),
  type = col_character(),
  organization = col_character()
)

assignee_tbl <- vroom(
  file       = "DS_101/02_data_wrangling/Patent_data_reduced/assignee.tsv", 
  delim      = "\t", 
  col_types  = col_types_assignee,
  na         = c("", "NA", "NULL")
)

# Patent assignee: ----

col_types_patent_assignee <- list(
  patent_id = col_character(),
  assignee_id = col_character()
)

patent_assignee_tbl <- vroom(
  file       = "DS_101/02_data_wrangling/Patent_data_reduced/patent_assignee.tsv", 
  delim      = "\t", 
  col_types  = col_types_patent_assignee,
  na         = c("", "NA", "NULL")
)

col_types_uspc <- list(
  patent_id = col_character(),
  mainclass_id = col_number(),
  sequence = col_number()
)


uspc_tbl <- vroom(
  file       = "DS_101/02_data_wrangling/Patent_data_reduced/uspc.tsv", 
  delim      = "\t", 
  col_types  = col_types_uspc,
  na         = c("", "NA", "NULL")
)


# 3.0 Converting Data Structure ----

setDT(assignee_tbl)
setDT(patent_tbl)
setDT(patent_assignee_tbl)
setDT(uspc_tbl)
```

## Which US company has the most patents?

```{r}
# 4.0 DATA WRANGLING ----
# Q1.What US company / corporation has the most patents? 

setnames(assignee_tbl, "id", "assignee_id")

combined_data <- merge(x = patent_assignee_tbl, y = assignee_tbl, by = "assignee_id")


us_patents <- combined_data %>%
  filter(type == 2)%>%
  filter(!is.na(patent_id) || !is.na(organization)) %>%
  select(-type, -assignee_id)%>% 
  group_by(organization) %>%
  count(patent_id) %>%
  select(-patent_id)%>%
  summarise(total = sum(n))%>%
  arrange(desc(total))   

us_top_10 <- us_patents %>% slice(1:10)
us_top_10
```

## Recent patent activity: What US company had the most patents granted in 2019?

```{r}
# Q2. What US company had the most patents granted in 2019? 


tbl_2 <- patent_tbl %>%   
         separate(col  = date,
         into = c("year", "month", "day"),
          sep  = "-", remove = TRUE) %>%
          mutate(
              month = as.numeric(month)
            )%>%
          filter(month == 01)%>%
          select(-year, -day)

setnames(tbl_2, "id", "patent_id")
combined_data_2 <- merge(x = tbl_2, y = combined_data, by = "patent_id")

us_top_10_2014 <- combined_data_2%>%
                    filter(type == 2)%>%
                    filter(!is.na(patent_id) || !is.na(organization)) %>%
                    select(organization, patent_id) %>%
                    group_by(organization) %>%
                    count(patent_id) %>%   
                    summarise(total_patents = sum(n))%>%
                    arrange(desc(total_patents)) %>% slice(1:10)  
us_top_10_2014

us_top_10_2014_new <- combined_data_2%>%
                        filter(type == 2 & num_claims == 1)%>%
                        filter(!is.na(patent_id) || !is.na(organization)) %>%
                        select(organization, patent_id) %>%
                        group_by(organization) %>%
                        count(patent_id) %>%   
                        summarise(total_patents = sum(n))%>%
                        arrange(desc(total_patents)) %>% slice(1:10)
us_top_10_2014_new

```

## What is the most innovative tech sector? For the top 10 companies (worldwide) with the most patents, what are the top 5 USPTO tech main classes?

```{r}
#Q3. What is the most innovative tech sector? 
#For the top 10 companies (worldwide) with the most patents,
#what are the top 5 USPTO tech main classes?

combined_data_3 <- merge(x = uspc_tbl, y = combined_data_2, by = "patent_id")



top10_worlwide_patents <- combined_data_3  %>%
                  filter(!is.na(patent_id) || !is.na(organization))%>%
                  group_by(organization) %>%
                  arrange(desc(mainclass_id)) %>%
  # set mainclass order first, the result will be sorted automatically 
                  count(patent_id) %>%
                  select(-patent_id)%>%
                  summarise(total_patents_wordwide = sum(n))%>%
                  ungroup() %>%
                  arrange(desc(total_patents_wordwide)) %>% slice(1:10)  

top10_worlwide_patents 
```

# Data Visualization (Chapter 5)

This is the last chapter and the conclusion of "Data Science Basics" course at TUHH. The objective of this chapter is to learn the proper tools for visualizing data so as to convert the data in table or text format to a graph from which meaningful info can be extracted.

## Challenge 1 Plot

This plot depicts the rate of change of COVID-19 cases along the year 2020, and determines which country has a significance in the number of cases.

```{r plot2, fig.width=8, fig.height=6}
library(scales)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(readxl)
library(ggthemes)
library(dplyr)

covid_data_tbl <- read_csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")

# Preparing Data for plotting
covid_data_wrangeled_tbl <- covid_data_tbl %>%
  select(countriesAndTerritories, cases, dateRep, month, year, day) %>%
  relocate(year, month, day) %>%
  filter(year == 2020, month >= 1) %>%
  filter(day != 1) %>%
  filter(countriesAndTerritories == "France" | countriesAndTerritories == "Germany" | countriesAndTerritories == "United_Kingdom" | countriesAndTerritories == "Spain" | countriesAndTerritories == "United_States_of_America") %>%
  group_by(countriesAndTerritories,month) %>%
  summarize(totalcases = sum(cases)) %>%
  ungroup()
    

covid_data_wrangeled_tbl %>%
  ggplot(aes(month ,totalcases, color = countriesAndTerritories)) +
  geom_smooth(method = "loess", span = 0.2) +
  scale_y_continuous(labels = scales::dollar_format(scale  = 1/1e6, 
                                                    prefix = "", 
                                                    suffix = "M")) +
  scale_x_continuous(breaks = seq(1, 11 , by=1),labels= c("January",
                                                          "February",
                                                          "March",
                                                          "April",
                                                          "May",
                                                          "June",
                                                          "July",
                                                          "August",
                                                          "September",
                                                          "October",
                                                          "November")) +

  labs(
    title = ("COVID-19 confirmed cases worldwide"),
    subtitle = ("United States has the highest rate of cases"),
    caption = "",
    x = "(Year 2020)",
    y = "Cumulative Cases",
    color = "Country"
      ) +
  geom_label(aes(label = (totalcases)), 
             hjust = "inward",
             size  = 3,
             color = RColorBrewer::brewer.pal(n = 12, name = "Blues")[8]) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Challenge 2 Plots

The goal of this challenge is to visualize the distribution of the mortality rate (deaths / population)

```{r code2 plot3, results='hide', fig.width=9, fig.height=6}
library(tidyverse)
library(scales)
library(ggplot2)
library(lubridate)
library(ggthemes)
library(dbplyr)
library(maps)

# Importing data


covid_data_tbl <- read_csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")

world <- map_data("world") %>%
  mutate(across(region, str_replace_all, "_", " ")) %>%
  mutate(region = case_when(
    
    region == "UK" ~ "United_Kingdom",
    region == "USA" ~ "United_States_of_America",
    region == "Czech_Republic" ~ "Czechia",
    TRUE ~ region
    
  ))

covid_data_tbl %>% 
  mutate(across(countriesAndTerritories, str_replace_all, "_", " ")) %>%
  mutate(countriesAndTerritories = case_when(
    
    countriesAndTerritories == "United Kingdom" ~ "United_Kingdom",
    countriesAndTerritories == "United_States_of_America" ~ "United States of America",
    countriesAndTerritories == "Czechia"~"Czechia",
    TRUE ~ countriesAndTerritories
    
  ))

#manipulation of world data table
world_map <- world %>%
  select(region, long, lat, group) %>%
  rename(countriesAndTerritories = region)


#manipulation of covid data table
covid_modified_data_tbl <- covid_data_tbl %>%
  select(day, month, year, countriesAndTerritories, deaths, popData2019) %>%
  group_by(year, countriesAndTerritories, popData2019) %>%
  summarise(total_death = sum(deaths)) %>%
  ungroup() %>%
  mutate(mortality_rate = (total_death / popData2019) * 100)

#merging data between 2 tables 
All_data_tbl <- left_join(covid_modified_data_tbl,
                          world_map,
                          by = "countriesAndTerritories") %>%
                filter(year == 2020)

#first layer of the map
world_map <- map_data("world")
ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill="lightblue", colour = "black",size = 0.1)

#second layer of the map
ggplot(data = All_data_tbl, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = mortality_rate), color = "blue",size = 0.1) +
  scale_fill_viridis_c(option = "E", alpha = 0.75 )
```